\chapter{Body-worn sensor technology}

\section{Sensor technology}
Basis technology (mention activPAL) mention gps for tracking of old people

\section{Presentation of sensor data}

\subsection{Nike+}
NikeFuel is a unit of measurement used by all Nike activity trackers, however there are no details on how activity is converted to NikeFuel, as the algorithm is proprietary. The wristband does calculate steps and calories burned, but the NikeFuel is the prime focus of their product line. NikeFuel does not take into account gender, height, weight, but looks purely at activity. Meaning that a mile will award the same amount of points to users of very different physiology. The daily progress is represented through a ring that fills up when the FuelBand detects activity, the ring is full once the daily activity goal is reached. If the NikeFuel goal is exceeded by a certain percentage the ring will gain visual improvements. 

\section{Personal informatics and quantified self}
Currently there are two names that stand out within self-monitoring: Quantified Self, and Personal Informatics. Quantified self is a community of end users who share data and exchange experiences with tools that help them collect information. Personal Informatics on the other hand is academic and developer driven, they desire to understand what makes a good tool, and how the user interacts with such technology.

\subsection{Quantified Self}
In 2011 a movement known as Quantified Self*\cite{quantifiedSelf} had their first conference \cite{bodyHackers}, here people shared data that they had collected about themselves width different types of devices. The goal is to gather as much information about yourself as possible, to learn about yourself through quantitative data. Members of Quantified Self* collect information about everything from sleep patterns and diets to mood and stress levels.

(Cut the next paragraph to shorten this section?)

To promote further development in tools that gather these types of information, the participants of Quantified Self has worked closely with companies and individuals that create personal informatics tools. Devices such as Nike's FuelBand \cite{fuelBand} and Fitbit \cite{fitBit} are results of this cooperation, and both products have been well received.

\subsection{Personal Informatics}
Personal informatics is the label used to classify tools that help people collect personal information for the purpose of self-monitoring and self reflection. These tools are used to help individuals gain self-knowledge about their behaviour, habits, and thoughts\cite{personalInformatics}.

(Cut the next paragraph to shorten the section?)

The Computer-Human Interaction (CHI) conference has since 2010 \cite{chi2010} held workshops and accepted papers on Personal Informatics. The aim is to increase understanding of how the tools affect the user, explore new possibilities, and overall improvement of the user experience.

\section{Medical applications (not a priority)}
Find articles that use body worn sensors for health care