\chapter{Focus Group 2}
\label{ch:focusGroup2}
The second focus group had several goals: Verify the scenarios and requirements created after the first focus group, review and get feedback on prototype 2 and interview the participants about a normal work day. Much like the previous focus group chapter this one starts with introducing the participants and the process, before presenting the results.

\section{Participants}
\label{sec:participants}
Participants from the previous focus group session were invited, but due to conflicts in the schedule only four out of the five could make it. For this session there were three females and one male present. The ages were between 34--36. Table~\ref{tab:participants} provides more detailed information about each participant.

The information gathered by the questionnaire showed that all the participants used laptops and smartphones on a daily basis. One of the participants also used tablets daily. All the participants were very positive to using new technology in their practice. They also agreed that new technology would improve the quality and effectiveness of their work.

\begin{table}[h!]
  \begin{center}
  \begin{tabular}{|c|c|c|c|}
    \hline
    \textbf{ID} & \textbf{Gender} & \textbf{Age} & \textbf{Position} \\ \hline
    P1 & Female & 35 & Physiotherapist \\ \hline
    P2 & Female & 35 & Physiotherapist \\ \hline
    P3 & Female & 34 & Physiotherapist \\ \hline
    P4 & Male   & 36 & Physiotherapist \\ \hline
    P5 & Female & 45 & Head of group   \\ \hline
  \end{tabular}
  \end{center}
  \caption[Participant information]{Basic information about the participants in the focus group.}
  \label{tab:participants}
\end{table}

One of the participants is referred to as \emph{head of group}, a translation of the Norwegian term ``fagleder''. 20\% of her work is with patients, the rest of her job involves increasing the knowledge and expertise of physiotherapists working with the elderly. She makes sure national plans and guidelines are incorporated into their daily work, and acts as an advisor in cases where additional assistance is required. She is also responsible for looking at new research and technology that could be used in physiotherapy.

\section{Location and Process}
The focus group was conducted at St. Olav's Hospital, but we were unable to obtain the exact same room.  Much like the first session the participants were seated around a table facing a projector canvas. The visualizations were controlled by the assistant moderator while the primary moderator was responsible for conducting the session. Like the previous session it was video recorded and lasted for two hours.

The agenda for the session:
\vspace{-6mm}
\begin{enumerate}[itemsep=0cm, parsep=0cm]
  \item Introduction and questionnaire
  \item Discuss scenarios
  \item Review of visualizations
  \item Review the requirements
  \item Discussion
  \item Interview about physiotherapy in practice
\end{enumerate}

The introduction phase was brief, since we were already acquainted with the participants. After the agenda was presented the participants were asked to fill out a questionnaire about their familiarity with technological aids such as smartphones and tablets, as well as how they felt about using information visualizations in their work. 

After the questionnaire, scenarios created after the first focus group were presented to the participants so that feedback could be received. They were asked to provide additional scenarios if they felt that the existing ones did not cover all the relevant uses of the technology presented.

The visualizations were presented one by one as in the first focus group. The participants then gave feedback on the modifications made as well as the new features. When all the visualizations had been presented and discussed, the participants were given different colour choices for visualization T1. Laptops were used for this purpose as it was hard to differentiate between the colour gradients on the projector. A printed black and white version of visualizations T1 and T5 were provided to the participants, while the rest were shown on the computer in a grayscale.

Requirements created based on the feedback received from the first focus group were then presented to the participants. After a brief introduction to software requirements, the participants were asked to go through them and comment if they felt something was missing or unclear.

After the requirements had been reviewed the findings of the session were discussed so that there were no ambiguities. The last part of the session was used for an unstructured interview with the physiotherapists about how physiotherapy is practised in Norway. 

\section{Results}
This section contains the results of the second focus group. New revisions of the scenarios and the requirements are presented, and the participant's feedback and comments on the visualizations in the second prototype are covered in some detail. The last two sections contain a discussion about colour choices and printouts of the visualizations, and a description of how physiotherapy is practices in Norway.

\subsection{Scenarios}
The scenarios created after focus group 1 were reviewed in the second focus group. After discussing the scenarios with the participants we created a new version, as seen in the Table~\ref{tab:scen2}:

\begin{table}[h!]
  \centering
  \begin{tabular}{|c|p{9cm}|}
    \hline
    \textbf{Id} & \textbf{Scenario} \\ \hline
    S2-1 & When analysing patients activity level, either individually or in cooperation with occupational therapists or other physiotherapists. \\ \hline
    S2-2 & In consultation with the patient. \\ \hline
    S2-3 & In communication with nursing homes and home care personnel. \\ \hline
    S2-4 & In consultation with next of kin. \\ \hline
    S2-5 & For educational purposes. \\ \hline
  \end{tabular}
  \caption[Scenarios after the second focus group]{Table showing the scenarios after the second focus group.}
  \label{tab:scen2}
\end{table}

S1-1 was changed to include occupational therapists. Occupational therapists are the colleagues that are most often consulted according to the physiotherapists. 

S1-2 was not changed. It is important to note that the physiotherapists estimated that maybe half of the patients would be cognitively capable of understanding the visualizations.

S1-3 was changed to include nursing homes as the most important partner to communicate with. Patients in nursing homes are in general much less active than those that live at home and there is therefore a greater need to inform the personnel about how inactive some of their patients are.

S1-4 was not changed.

S2-5 was a new scenario added on the request of the focus group. An increasing part of their work consists of tutoring home care personnel about exercises that patients can perform to increase their activity. The participants felt using visualizations in this setting would be useful. 

\subsection{Requirements}
The new revision of functional requirements can be seen in Table~\ref{tab:reqFoc2}. One requirement which all of the participants wished to clarify was R1-1. They were specific about the fact that it should not be national goals, but national recommendations, this was rectified in the new requirements. Two of the participants mentioned that an overview of the week was good, but receiving a summary of the daily activity was equally important. This was provided by the pie chart but was never explicitly stated as a requirement, R2-15 was added to reflect this.

Being able to include or exclude certain hours of the night was also discussed due it being brought up during several visualizations, R2-16 has been added as a response to this. There was disagreement on what counted as ``night'', and if a certain time interval should be set by the user or a pre-set value should be used. 

One of the things the participants felt was useful with visualizations F3 was the ability to detect how short the intervals of walking activity were, no other visualizations gave them this ability. R2-3 was changed to include the interval of any type of activity, not just inactivity. 

\begin{table}[h!]
  \begin{center}
  \begin{tabular}{|c|p{12cm}|}
    \hline
      \textbf{Id} & \textbf{Requirement} \\ \hline
    \multicolumn{2}{|l|}{The visualizations should \ldots} \\ \hline
      R2-1 & give the user an overview of the week where the days are classified by national recommendations or personal goals \\ \hline
      R2-2 & show the activity level for each hour of the day \\ \hline
      R2-3 & make it easy to identify the length of activity intervals \\ \hline
      R2-4 & make it possible to compare multiple days \\ \hline
      R2-5 & make it easy to identify hours of the day where activity can be added \\ \hline
      R2-6 & show the activity level compared to national or personal goals \\ \hline
      R2-7 & let the user identify patients that are active during the night \\ \hline
      R2-8 & let the user compare two separate weeks to see the patients progress \\ \hline
      R2-9 & should be printable in grayscale \\ \hline
      R2-15 & show the activity distribution for a day (sedentary, standing, walking) \\ \hline
      R2-16 & allow the users to toggle if nighttime should be included or not \\ \hline
  \end{tabular}
  \end{center}
  \caption[Functional requirements after the second focus group]{Functional requirements from the second focus group.}
  \label{tab:reqFoc2}
\end{table}

When the participants were presented with the user experience requirements they were initially a bit confused of what exactly these requirements entailed. Some time was used to explain the meaning of UX requirements to the participants. The revised UX requirements can be found in Table~\ref{tab:uxReqFoc2}. R1-13 was clarified to include not just the user, but other non medical partners as well, this is now reflected in R2-13. According to the participants some patients are simply not well enough cognitively to understand the visualizations no matter how simple they are, R1-14 was therefore changed accordingly.

\begin{table}[h!]
  \begin{center}
  \begin{tabular}{|c|p{12cm}|}
    \hline
      \textbf{Id} & \textbf{Requirement} \\ \hline
    \multicolumn{2}{|l|}{The visualizations should \ldots} \\ \hline
      R2-10 & not be judgemental towards the patients activity level \\ \hline
      R2-11 & should be honest about the patients activity level \\ \hline
      R2-12 & should motivate the patient to be more active \\ \hline
      R2-13 & should be intuitive and easy to understand for the user and third parties \\ \hline
      R2-14 & be easy to explain to cognitively capable patients \\ \hline
  \end{tabular}
  \end{center}
  \caption[User experience requirements after the second focus group]{User experience requirements from the second focus group.}
  \label{tab:uxReqFoc2}
\end{table}

\subsection{Visualizations}
The second prototype contained six visualizations, where five of them were modified versions of visualizations present in the first prototype and one was new. Other features that were added were the ability to set goals and a small chart in the form of circles for displaying how far the patient was from reaching his or her goal.

\subsubsection{Overview chart}
U1 was the only overview chart left after changes were made for the second prototype. Participants of the focus group were pleased with the fact that the smilies had been replaced with coloured squares. The new way of classifying days based on how close they were to their set goals was well received by all the participants. One of the participants stated that it was important to be able to set custom goals, because how active patients should be differs greatly. Another participant suggested adding the current goal to the visualization to make it easier to determine what each classification means. Most of the participants liked the simplicity of U1 and said that this was a type of diagram that many of their patients would understand, however one of the participants felt that the lack of detail in the diagram made it useless in practical situations.

Most of the participants like the new overview chart. The currently set goal should be displayed so it is clear how the days are being classified.

\subsubsection{Aggregated charts}
F2, the box diagram, was discarded in the second prototype, but the illustrations used to show each type of activity was added to F1. Nighttime was also removed, as requested in the first focus group. The participants were satisfied with the changes and liked the fact that the illustrations were now added to the pie chart, F1. All the participants agreed that the F1 was a good way to see the overall activity of the day, and that excluding nighttime increased the practical use of this chart.

F3 was not changed other than the colour for sedentary activity being white instead of red. The participants did not like the colour change particularly, and felt that the red would be better than white. All the participants agreed that this was a good way to visualize interval length. Some of the participants suggested that one should have the ability toggle nighttime on and off. They stated that nighttime is relevant for some patients that are active during the night, but for most patients displaying nighttime would just be in the way.

Both F1 and F3 were well received by the physiotherapists. F3 should have functionality for toggling nighttime on and off.

\subsubsection{Timeline and Clock}
T2 and T3 were discarded after focus group 1, because the participants felt the amount of detail was unnecessary, as well as the active periods being hard to identify. T1 was the favourite visualization from focus group 1 and was not changed much for the second prototype. Goal circle diagrams were added to each day to display how the patient's activity level compared to the goal set. The participants did not like the addition of the goal circles. One suggested being able to toggle the goal circles on and off, or be able to look at the goal circles as a separate visualization. As in U1 the participants felt that the current value of the goal should be displayed, so that it is clear what a circle represents.

T4 was not particularly well received in the first focus group, but it was added to the second prototype with a colour change of the sedentary activity from red to white. Even with the colour change the participants all agreed that this type of chart was too detailed and was not useful in practice. The visualization should be discarded.

T5 was the only new visualization added for prototype 2, if not counting the goal circles. Though this visualization was made on the request of the participants, they were not overly enthusiastic about T5. Most of the participants preferred T1 to T5. However when asked if they would like the option of having both visualizations available they all agreed that T5 could be useful in some cases and should be kept. Also here the participants felt that the goal circles made the chart overly complex, and that this should be a toggle option.

T4 should be removed as it has no practical use. T1 and T5 should be kept, but the user should be able to toggle the goal circles on and off. The current goal should be shown when the goal circles are displayed. Goal circles should be its own visualization.

\subsection{Colours and Printouts}
Different colour choices were discussed with the participants. Because the projector did not handle gradients well, a laptop was used to show different colour choices. For the gradient colour the participants preferred white to blue to black, and white to black. For representing the different activity types (sitting, standing, walking) the participants liked red, yellow and green. One participant expressed some concern with the fact that red, yellow and green was used for sitting, standing and walking for F1, F3 and T5, but in U1 it was used for classifications. Using the same colours with different meanings can be confusing. An alternative to the coloured boxes in U1 should be found.

Because physiotherapists working for Trondheim Kommune do not have access to a colour printer, grayscale printouts were showed to the participants. The participants found it hard to use T1 effectively when printed, because the printer did not handle grayscale gradients well. In this situation the participants preferred T5 over T1 since bars were used instead of colour for displaying the activity level. U1 and F1 was found to work well on printouts, however F3 was less useful when printed because the functionality to hold over a ball for more information about the interval is lost.

\subsection{Physiotherapy in Practice} 
\label{sec:physiotherapyPractice}
To get an overview of how physiotherapy is conducted in Norway we performed an interview with the participants of the focus group. The physiotherapists that were recruited for the focus groups work with elderly and patients that otherwise needs to be helped from home (e.g. patients suffering from Cerebral Palsy). The seven step process below describes what happens when a patient is referred to a physiotherapist:

\vspace{-4mm}
\begin{enumerate}
  \item A general practitioner or other healthcare personnel can fill out an application for their patients to see a physiotherapist.
  \item The application is then evaluated and placed into a priority queue. Applications may be prioritized if the matter is time sensitive, such as recovery after a fracture or surgery.
  \item When a physiotherapist is available they are given the application on top of the priority queue.
  \item The physiotherapist then makes a house visit to the patient so that they may get an understanding of the current activity level. The activity level is mapped through several different types of exercises and through conversations with the patient. The checklist that physiotherapists use is provided in Appendix~\ref{checklist}. Overall posture and the speed of movements help assess the general state of the patient.
  \item The next step is to create an exercise plan for the patient in order to increase the activity level. During conversations with the patient the physiotherapist discusses what kind of improvements are realistic to achieve considering the current activity level, motivation, physical health, etc. All of this data is then used to create an exercise plan that the patient can follow to reach his goals.
  \item When an appropriate plan has been created the physiotherapist returns to the patient to explain how the exercises are executed as well as motivating the patient to reach his goals. The plan may also include other types of activity, for many patients something as simple as walking to the store to buy groceries can be enough to make a difference in the overall activity.
  \item Physiotherapist will return regularly to check up on the patient. The interval between check-ups will vary in respect to how well the physiotherapists expects the patient to follow the agreed upon plan. Some patients lack motivation, and will need more regular check-ups. During such meetings the physiotherapist will get an idea of how much the patient has improved. Emphasizing on the improvement made is an important factor in motivating the patient to stick to the plan. If there has been little or no change in the activity level, the physiotherapist may want to make changes to the plan so better results can be achieved.
\end{enumerate}
